Ecological and evolutionary implications of starvation and body size

### 2016-07-25 Manuscript/2.tex
Sid first pass

%Extinction to the left and the right
\paragraph{Extinction risk}
Within our model higher rates of starvation result in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality. However, from the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}). We examine the competing effects of cyclic dynamics vs. changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.
We computed the probability of extinction, where extinction is defined as $H(t)+F(t)=10$ at any instant across all values of $10^2<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq. \ref{eq:system} for an organism of $M=100$ grams, assuming random initial conditions around the steady state (Eq. \ref{eq:ss}). %random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
By allowing the rate of starvation to vary, we assessed extinction risk across a range of values of the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a horizontal cross-section of Fig. \ref{fig:fp}.
As expected, higher rates of extinction correlated with both low and high values of $\sigma/\rho$; for low values the higher extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig. \ref{fig:ext}).
For large values of $\sigma/\rho$, higher extinction risk is due to the decrease in the steady state consumer population density.
This interplay creates an `extinction refuge' as shown in Fig. \ref{fig:ext}, such that for a relatively constrained range of $\sigma/\rho$, extinction probabilities are minimized.

>>>>

%Extinction to the left and the right
\paragraph{Extinction risk}
Within our model higher rates of starvation result in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality. However, from the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}). We examine the competing effects of cyclic dynamics vs. changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.
We computed the probability of extinction, where extinction is assumed if the population trajectory goes below $0.2~\times$ the allometrically constrained steady state across all values of $10^2<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq. \ref{eq:system} for an organism of $M=100$ grams, assuming random initial conditions around the steady state (Eq. \ref{eq:ss}). %random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
By allowing the rate of starvation to vary, we assessed extinction risk across a range of values of the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a horizontal cross-section of Fig. \ref{fig:fp}.
As expected, higher rates of extinction correlated with both low and high values of $\sigma/\rho$; for low values the higher extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig. \ref{fig:ext}).
For large values of $\sigma/\rho$, higher extinction risk is due to the decrease in the steady state consumer population density.
This interplay creates an `extinction refuge' as shown in Fig. \ref{fig:ext}, such that for a relatively constrained range of $\sigma/\rho$, extinction probabilities are minimized.
